---
title: Normalization and Variance Modelling
---

```{r}
library(scater)
library(scran)
library(BiocStyle)
library(pheatmap)
sce.grun.hsc <- readRDS("output/grun_quality_control_sce.RDS")
```

# Normalization

```{r normalization}
library(scran)
set.seed(101000110)
clusters <- quickCluster(sce.grun.hsc)
sce.grun.hsc <- computeSumFactors(sce.grun.hsc, clusters=clusters)
sce.grun.hsc <- logNormCounts(sce.grun.hsc)
```

We examine some key metrics for the distribution of size factors, and compare it to the library sizes as a sanity check.

```{r}
summary(sizeFactors(sce.grun.hsc))
```

```{r unref-hgrun-norm, fig.cap="Relationship between the library size factors and the deconvolution size factors in the Grun HSC dataset."}
plot(librarySizeFactors(sce.grun.hsc), sizeFactors(sce.grun.hsc), pch=16,
    xlab="Library size factors", ylab="Deconvolution factors", log="xy")
```

# Variance modelling

We create a mean-variance trend based on the expectation that UMI counts have Poisson technical noise.
We do not block on sample here as we want to preserve any difference between the micro-dissected cells and the sorted HSCs.

```{r variance-modelling}
set.seed(00010101)
dec.grun.hsc <- modelGeneVarByPoisson(sce.grun.hsc) 
top.grun.hsc <- getTopHVGs(dec.grun.hsc, prop=0.1)
```

The lack of a typical "bump" shape in Figure \@ref(fig:unref-hgrun-var) is caused by the low counts.

```{r unref-hgrun-var, fig.cap="Per-gene variance as a function of the mean for the log-expression values in the Grun HSC dataset. Each point represents a gene (black) with the mean-variance trend (blue) fitted to the simulated Poisson-distributed noise."}
plot(dec.grun.hsc$mean, dec.grun.hsc$total, pch=16, cex=0.5,
    xlab="Mean of log-expression", ylab="Variance of log-expression")
curfit <- metadata(dec.grun.hsc)
curve(curfit$trend(x), col='dodgerblue', add=TRUE, lwd=2)
```

```{r export}
saveRDS(sce.grun.hsc,"output/grun_normalization_sce.RDS")
saveRDS(dec.grun.hsc,"output/grun_normalization_dec.RDS")
saveRDS(top.grun.hsc,"output/grun_normalization_top.RDS")
```
