---
title: Quality Control
---

```{r}
sce.grun.hsc <- readRDS("output/grun_import_sce.RDS")
```


```{r}
unfiltered <- sce.grun.hsc
```

For some reason, no mitochondrial transcripts are available, and we have no spike-in transcripts, so we only use the number of detected genes and the library size for quality control.
We block on the protocol used for cell extraction, ignoring the micro-dissected cells when computing this threshold.
This is based on our judgement that a majority of micro-dissected plates consist of a majority of low-quality cells, compromising the assumptions of outlier detection.

```{r quality-control}
library(scuttle)
stats <- perCellQCMetrics(sce.grun.hsc)
qc <- quickPerCellQC(stats, batch=sce.grun.hsc$protocol,
    subset=grepl("sorted", sce.grun.hsc$protocol))
sce.grun.hsc <- sce.grun.hsc[,!qc$discard]
```

We examine the number of cells discarded for each reason.

```{r}
colSums(as.matrix(qc))
```

We create some diagnostic plots for each metric.
The library sizes are unusually low for many plates of micro-dissected cells; this may be attributable to damage induced by the extraction protocol compared to cell sorting.

```{r unref-hgrun-qc-dist, fig.asp=0.8, fig.wide=10, fig.cap="Distribution of each QC metric across cells in the Grun HSC dataset. Each point represents a cell and is colored according to whether that cell was discarded."}
colData(unfiltered) <- cbind(colData(unfiltered), stats)
unfiltered$discard <- qc$discard

library(scater)
gridExtra::grid.arrange(
    plotColData(unfiltered, y="sum", x="sample", colour_by="discard", 
        other_fields="protocol") + scale_y_log10() + ggtitle("Total count") +
        facet_wrap(~protocol),
    plotColData(unfiltered, y="detected", x="sample", colour_by="discard",
        other_fields="protocol") + scale_y_log10() + 
        ggtitle("Detected features") + facet_wrap(~protocol),
    ncol=1
)
```

```{r}
saveRDS(sce.grun.hsc,"output/grun_quality_control_sce.RDS")
```

