# Setup

## Data Import

Reading in 'iris' data exported from the [generate](generate.html) page.

```{r}
iris = read.csv("output/iris.csv")
```

## PCA Parameters

Using an analysis configuration file (`code/analysis_config.txt`) to control some parameters. Using the following species:

```{r}
ana_species = readLines("code/analysis_config.txt")
ana_species
```

Tabulating rows remaining for each species:

```{r}
# remove the first column
iris = iris[, -1]

# leave only the species to be analysed
iris = iris[iris$Species %in% ana_species, ]

table(iris$Species)
```

# PCA

Using the first 4 columns for principal component analysis.

```{r}
mat = iris[, c(1,2,3,4)]
head(iris)
```

Calculating principal component scores.

```{r}
# calculate principal compomnents for 'iris' dataset
iris_pca = prcomp(mat,scale.=TRUE)

# save all PCs, along with the species names
iris_pca = data.frame(iris_pca$x)
iris_pca = cbind(iris_pca,iris)
```

## PCA Results

The set of scores was combined with the original data to obtain this data.frame:

```{r}
print(iris_pca[1:10, ])
```

### Score Pair Visualization

Plotting pairs of principal components with `pairs()`.

```{r}
species = iris_pca[, 9]

# plot
pairs(iris_pca[,1:4], col=as.factor(species), main="iris")
```

Comparing this to the original columns.

```{r}
pairs(iris_pca[,5:8],col=as.factor(species))
```

## Export PC Scores

The results will be exported for downstream statistical analysis.

```{r}
# write to 'output/iris_pca.csv'
write.csv(iris_pca, file="output/iris_pca.csv")
```

