---
title: Further Exploration
---

```{r}
source("src/env.R")
sce.nest <- readRDS("output/nestorowa_normalization_sce.RDS")
top.nest <- readRDS("output/nestorowa_normalization_top.RDS")
dec.nest <- readRDS("output/nestorowa_normalization_dec.RDS")
```

# Dimensionality reduction

```{r dimensionality-reduction}
set.seed(101010011)
sce.nest <- denoisePCA(sce.nest, technical=dec.nest, subset.row=top.nest)
sce.nest <- runTSNE(sce.nest, dimred="PCA")
```

We check that the number of retained PCs is sensible.

```{r}
ncol(reducedDim(sce.nest, "PCA"))
```

# Clustering

```{r clustering}
snn.gr <- buildSNNGraph(sce.nest, use.dimred="PCA")
colLabels(sce.nest) <- factor(igraph::cluster_walktrap(snn.gr)$membership)
```

```{r}
table(colLabels(sce.nest))
```

```{r unref-nest-tsne, fig.cap="Obligatory $t$-SNE plot of the Nestorowa HSC dataset, where each point represents a cell and is colored according to the assigned cluster."}
plotTSNE(sce.nest, colour_by="label")
```

# Marker gene detection

```{r marker-detection}
markers <- findMarkers(sce.nest, colLabels(sce.nest), 
    test.type="wilcox", direction="up", lfc=0.5,
    row.data=rowData(sce.nest)[,"SYMBOL",drop=FALSE])
```

```{r, echo=FALSE}
cluster <- "8"
```

To illustrate the manual annotation process, we examine the marker genes for one of the clusters.
Upregulation of _Car2_, _Hebp1_ amd hemoglobins indicates that cluster `r cluster` contains erythroid precursors.

```{r unref-heat-nest-markers, fig.asp=1.5, fig.cap=sprintf("Heatmap of the AUCs for the top marker genes in cluster %s compared to all other clusters.", cluster)}
chosen <- markers[['8']]
best <- chosen[chosen$Top <= 10,]
aucs <- getMarkerEffects(best, prefix="AUC")
rownames(aucs) <- best$SYMBOL

library(pheatmap)
pheatmap(aucs, color=viridis::plasma(100))
```

```{r, echo=FALSE}
# Checking my statements:
stopifnot(identical(chosen, markers[[cluster]]))
stopifnot(all(c("Car2", "Hba-a2", "Hba-a1", "Hebp1") %in% rownames(aucs)))
```

# Cell type annotation

```{r cell-annotation}
library(SingleR)
mm.ref <- MouseRNAseqData()

# Renaming to symbols to match with reference row names.
renamed <- sce.nest
rownames(renamed) <- uniquifyFeatureNames(rownames(renamed),
    rowData(sce.nest)$SYMBOL)
labels <- SingleR(renamed, mm.ref, labels=mm.ref$label.fine)
```

Most clusters are not assigned to any single lineage (Figure \@ref(fig:unref-assignments-nest)), which is perhaps unsurprising given that HSCs are quite different from their terminal fates.
Cluster `r cluster` is considered to contain erythrocytes, which is roughly consistent with our conclusions from the marker gene analysis above.

```{r unref-assignments-nest, fig.cap="Heatmap of the distribution of cells for each cluster in the Nestorowa HSC dataset, based on their assignment to each label in the mouse RNA-seq references from the _SingleR_ package."}
tab <- table(labels$labels, colLabels(sce.nest))
pheatmap(log10(tab+10), color=viridis::viridis(100))
```

```{r, echo=FALSE}
test <- tab[,"8"]
stopifnot(test["Erythrocytes"]/sum(test) > 0.9)
```

# Miscellaneous analyses

This dataset also contains information about the protein abundances in each cell from FACS.
There is barely any heterogeneity in the chosen markers across the clusters (Figure \@ref(fig:unref-nest-facs));
this is perhaps unsurprising given that all cells should be HSCs of some sort.

```{r unref-nest-facs, fig.cap="Heatmap of the centered log-average intensity for each target protein quantified by FACS in the Nestorowa HSC dataset."}
Y <- colData(sce.nest)$FACS
keep <- rowSums(is.na(Y))==0 # Removing NA intensities.

se.averaged <- sumCountsAcrossCells(t(Y[keep,]), 
    colLabels(sce.nest)[keep], average=TRUE)
averaged <- assay(se.averaged)

log.intensities <- log2(averaged+1)
centered <- log.intensities - rowMeans(log.intensities)
pheatmap(centered, breaks=seq(-1, 1, length.out=101))
```

```{r, echo=FALSE}
# Check that my trash talk is correct.
rr <- rowRanges(centered)
stopifnot(all(rr[,2] - rr[,1] < 1))
```

