---
title: Further Exploration
---

```{r}
library(scater)
library(scran)
library(BiocStyle)
library(pheatmap)
sce.paul <- readRDS("output/paul_normalization_sce.RDS")
dec.paul <- readRDS("output/paul_normalization_dec.RDS")
top.paul <- readRDS("output/paul_normalization_top.RDS")
```

# Dimensionality reduction

```{r dimensionality-reduction}
set.seed(101010011)
sce.paul <- denoisePCA(sce.paul, technical=dec.paul, subset.row=top.paul)
sce.paul <- runTSNE(sce.paul, dimred="PCA")
```

We check that the number of retained PCs is sensible.

```{r}
ncol(reducedDim(sce.paul, "PCA"))
```

# Clustering

```{r clustering}
snn.gr <- buildSNNGraph(sce.paul, use.dimred="PCA", type="jaccard")
colLabels(sce.paul) <- factor(igraph::cluster_louvain(snn.gr)$membership)
```

These is a strong relationship between the cluster and the experimental treatment, which is to be expected.
Of course, this may also be attributable to some batch effect; the confounded nature of the experimental design makes it difficult to make any confident statements either way.

```{r unref-paul-heat, fig.cap="Heatmap of the distribution of cells across clusters (rows) for each experimental treatment (column)."}
tab <- table(colLabels(sce.paul), sce.paul$Batch_desc)
rownames(tab) <- paste("Cluster", rownames(tab))
pheatmap::pheatmap(log10(tab+10), color=viridis::viridis(100))
```

```{r unref-paul-tsne, fig.cap="Obligatory $t$-SNE plot of the Paul HSC dataset, where each point represents a cell and is colored according to the assigned cluster."}
plotTSNE(sce.paul, colour_by="label")
```

```{r unref-paul-tsne2, fig.width=8, fig.height=8, fig.cap="Obligatory $t$-SNE plot of the Paul HSC dataset faceted by the treatment condition, where each point represents a cell and is colored according to the assigned cluster."}
plotTSNE(sce.paul, colour_by="label", other_fields="Batch_desc") + 
    facet_wrap(~Batch_desc)
```

