---
title: Normalization and Variance Modelling
---

```{r}
source("src/env.R")
sce.paul <- readRDS("output/paul_quality_control_sce.RDS")
```

# Normalization

```{r normalization}
library(scran)
set.seed(101000110)
clusters <- quickCluster(sce.paul)
sce.paul <- computeSumFactors(sce.paul, clusters=clusters)
sce.paul <- logNormCounts(sce.paul)
```

We examine some key metrics for the distribution of size factors, and compare it to the library sizes as a sanity check.

```{r}
summary(sizeFactors(sce.paul))
```

```{r unref-paul-norm, fig.cap="Relationship between the library size factors and the deconvolution size factors in the Paul HSC dataset."}
plot(librarySizeFactors(sce.paul), sizeFactors(sce.paul), pch=16,
    xlab="Library size factors", ylab="Deconvolution factors", log="xy")
```

# Variance modelling

We fit a mean-variance trend to the endogenous genes to detect highly variable genes.
Unfortunately, the plates are confounded with an experimental treatment (`Batch_desc`) so we cannot block on the plate of origin.

```{r variance-modelling}
set.seed(00010101)
dec.paul <- modelGeneVarByPoisson(sce.paul)
top.paul <- getTopHVGs(dec.paul, prop=0.1)
```

```{r unref-paul-var, fig.cap="Per-gene variance as a function of the mean for the log-expression values in the Paul HSC dataset. Each point represents a gene (black) with the mean-variance trend (blue) fitted to simulated Poisson noise."}
plot(dec.paul$mean, dec.paul$total, pch=16, cex=0.5, 
    xlab="Mean of log-expression", ylab="Variance of log-expression")
curve(metadata(dec.paul)$trend(x), col="blue", add=TRUE)
```

```{r export}
saveRDS(sce.paul,"output/paul_normalization_sce.RDS")
saveRDS(dec.paul,"output/paul_normalization_dec.RDS")
saveRDS(top.paul,"output/paul_normalization_top.RDS")
```

