---
title: Quality Control
---

```{r}
sce.paul <- readRDS("output/paul_import_sce.RDS")
```

```{r}
unfiltered <- sce.paul
```

For some reason, only one mitochondrial transcripts are available, so we will perform quality control using only the library size and number of detected features.
Ideally, we would simply block on the plate of origin to account for differences in processing, but unfortunately, it seems that many plates have a large proportion (if not outright majority) of cells with poor values for both metrics.
We identify such plates based on the presence of very low outlier thresholds, for some arbitrary definition of "low"; we then redefine thresholds using information from the other (presumably high-quality) plates.

```{r quality-control}
library(scater)
stats <- perCellQCMetrics(sce.paul) 
qc <- quickPerCellQC(stats, batch=sce.paul$Plate_ID)

# Detecting batches with unusually low threshold values.
lib.thresholds <- attr(qc$low_lib_size, "thresholds")["lower",]
nfeat.thresholds <- attr(qc$low_n_features, "thresholds")["lower",]
ignore <- union(names(lib.thresholds)[lib.thresholds < 100],
    names(nfeat.thresholds)[nfeat.thresholds < 100])

# Repeating the QC using only the "high-quality" batches.
qc2 <- quickPerCellQC(stats, batch=sce.paul$Plate_ID,
    subset=!sce.paul$Plate_ID %in% ignore)
sce.paul <- sce.paul[,!qc2$discard]
```

We examine the number of cells discarded for each reason.

```{r}
colSums(as.matrix(qc2))
```

We create some diagnostic plots for each metric (Figure \@ref(fig:unref-paul-qc-dist)).

```{r unref-paul-qc-dist, fig.wide=TRUE, fig.cap="Distribution of each QC metric across cells in the Paul HSC dataset. Each point represents a cell and is colored according to whether that cell was discarded."}
colData(unfiltered) <- cbind(colData(unfiltered), stats)
unfiltered$discard <- qc2$discard
unfiltered$Plate_ID <- factor(unfiltered$Plate_ID)

gridExtra::grid.arrange(
    plotColData(unfiltered, y="sum", x="Plate_ID", colour_by="discard") +
        scale_y_log10() + ggtitle("Total count"),
    plotColData(unfiltered, y="detected", x="Plate_ID", colour_by="discard") +
        scale_y_log10() + ggtitle("Detected features"),
    ncol=1
)
```

```{r export}
saveRDS(sce.paul,"output/paul_quality_control_sce.RDS")
```

